---
title: "census_data_importclean"
output: html_document
---


### 1. Working with YH population data
Download Yorkshire & Humber LSOA population data from the [ONS](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/censusoutputareaestimatesintheyorkshireandthehumberregionofengland ) I chose mid-2015 for now

### 2. Add up the total population for each LSOA
The ONS data is categorised on LSOA11 and OA11 which means that there are several rows for each LSOA which need to be added together.
``` {r}
censuswhole2015 <- read_excel("mid2015census.xlsx", 
    sheet = "Mid-2015 Persons", skip = 4)
censuswhole2015 <- censuswhole2015 %>% rename(lsoa11 = LSOA11CD) %>% rename(totpop = "All Ages") 
censuswhole2015 = subset (censuswhole2015, select = c( "lsoa11", "totpop"))
censuswhole2015 <- censuswhole2015 %>% group_by (lsoa11) 

```

### 3. Count stratas in census data
#Download Yorkshire & Humber LSOA population data - I chose mid-2015
### https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/populationestimates/datasets/censusoutputareaestimatesintheyorkshireandthehumberregionofengland #
#in excel, for each sex, add in column 1=male, 2 = female, and change LSOA11CD heading to lsoa11#
#import each sheet to  R separately#

``` {r}
yhlsoamale <- read_excel("D:/DEPED/Working data/YH lsoa population.xls", sheet = "Mid-2015 Males")

yhlsoafemale <- read_excel("D:/DEPED/Working data/YH lsoa population.xls", 
                           sheet = "Mid-2015 Females")

#merge the male and female gen pops, then clean environment
yhlsoagenpop <- rbind(yhlsoamale, yhlsoafemale)

# count the number of people in age group for the yorkshire and humber general pop


## experimenting with 10 year age brackets
yhagegroup <- yhlsoagenpop %>%
  mutate(agegr1 = rowSums(.[4:13], na.rm=TRUE))  %>%
  mutate(agegr2 = rowSums(.[14:23], na.rm=TRUE)) %>%
  mutate(agegr3 = rowSums(.[24:33], na.rm=TRUE)) %>%
  mutate(agegr4 = rowSums(.[34:43], na.rm=TRUE)) %>%
  mutate(agegr5 = rowSums(.[44:54], na.rm=TRUE)) %>%
  mutate(agegr6 = rowSums(.[54:64], na.rm=TRUE)) %>%
  mutate(agegr7 = rowSums(.[64:74], na.rm=TRUE)) %>%
  mutate(agegr8 = rowSums(.[74:94], na.rm=TRUE)) %>%
  mutate(totpop = rowSums(.[4:94], na.rm=TRUE))
yhagegroup = subset (yhagegroup, select = c( "sex", "lsoa11","agegr1", "agegr2", "agegr3", "agegr4", "agegr5", "agegr6", "agegr7", "agegr8", "totpop"))
yhagegroup <- yhagegroup %>% group_by (sex, lsoa11) %>% summarise_all(funs(sum))


#add together multiple lsoa rows
yhagegroup <- yhagegroup %>% group_by (sex, lsoa11)

#clean the environment#
rm (yhlsoagenpop)



````

```{r}
yhagegroup <- yhagegroup %>% pivot_wider (names_from = c(sex), values_from = c (agegr1, agegr2, agegr3, agegr4, agegr5, agegr6, agegr7, agegr8, totpop ))
                                            
```
