---
title: "Bringing together code"
author: "Joanna Sutton-Klein"
output: html_document
---
<!--- rmarkdown::render("MasterR.Rmd") --->
rmarkdown::render("D:/DEPED/Github/edattendances2/MasterR.Rmd") 

This is the code that I am using to explore spatial variability in emergency department attendances.

<!--- Set working directory --->
```{r, echo=FALSE, cache=TRUE }
setwd("D:/DEPED/Github/edattendances2/")
```

<!--- Setting up R studio for the analysis --->
```{r, child = "Setting_up_Rstudio.Rmd", echo=FALSE, cache=TRUE}

```


The first stage is importing and cleaning the health data. In this section I also calculate the proportion of low acuity attendances for each locality

```{r, child = "HES_data_importclean.Rmd", echo=FALSE, cache=TRUE}

```

Next I downloaded the census data and calculate the total population for each LSOA

```{r, child = "census_data_importclean.Rmd", echo=FALSE, cache=TRUE}

```

Merge the health data and census data

```{r, child = "Merging_data.Rmd", echo=FALSE, cache=TRUE}

```

Calculate crude rates of attendance

```{r, child = "Calculating_rates.Rmd", echo=FALSE, cache=TRUE}

```

Calculate standardised rates of attendance

```{r, child = "standardisedrates.Rmd", echo=FALSE, cache=TRUE}

```

Maps of weighted rates

```{r, child = "mapweightedrate.Rmd", echo=FALSE, cache=TRUE}

```

